% Instructions: you don't need to change anything in the macros, but
% feel free to define new commands as you wish. Starting from the main
% body, change the specs (e.g., your
% name). use \begin{solution} \end{solution} environment to write your
% solutions. Don't forget to list your collaborators.

\documentclass[12pt,answers]{exam}
%============Macros==================%
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
%--------------Cosmetic----------------%
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{fullpage}
\usepackage{microtype}
\usepackage{xspace}
\usepackage[svgnames]{xcolor}
\usepackage[sc]{mathpazo}
\usepackage{enumitem}
\setlist[enumerate]{itemsep=1pt,topsep=2pt}
\setlist[itemize]{itemsep=1pt,topsep=2pt}
%----------Header--------------------%
\def\course{CS 584/684 Algorithm Design and Analysis}
\def\term{Portland State U, Winter 2021}
\def\prof{Lecturer: Fang Song}
\newcommand{\handout}[5]{
   \renewcommand{\thepage}{\arabic{page}}
   \begin{center}
   \framebox{
      \vbox{
    \hbox to 5.78in { \hfill \large{\course} \hfill }
       \vspace{2mm}
       \hbox to 5.78in { {\Large \hfill \textbf{#5}  \hfill} }
       \vspace{2mm}
       \hbox to 5.78in { \term \hfill \emph{#2}}
       \hbox to 5.78in { {#3 \hfill \emph{#4}}}
      }
   }
   \end{center}
   \vspace*{4mm}
}
\newcommand{\hw}[4]{\handout{#1}{#2}{#3}{#4}{Homework #1}}

%-----defs and commands-----%
\def\veps{\varepsilon}
\newcommand{\bit}{\{0,1\}}
\newcommand{\negl}{\text{negl}}
\newcommand{\corr}[1]{{\color{blue}{#1}}}
\newcommand{\alg}[1]{\textsf{#1}}
%=======Main document==============%
\begin{document}

%----Specs: change accordingly-----%
\newif\ifstudent % comment out false
\studenttrue 
%\studentfalse

\def\texbp{5} % bonus for typing in latex
\def\hwnum{6} %
\def\issuedate{02/18/21} % 
\def\duedate{03/02/21} % 
\def\yourname{your name} % type your name here

%------------------------------%
\ifstudent
\hw{\hwnum}{\issuedate}{Student: \yourname}{Due: \duedate}%
\else
\hw{\hwnum}{\issuedate}{\prof}{Due: \duedate}%
\fi

\noindent \textbf{Instructions.} This problem set contains \numpages\
pages (including this cover page) and \numquestions\ questions. A
random subset of problems will be graded.

\begin{itemize}
\item Your solutions will be graded on \emph{correctness} and
  \emph{clarity}. You should only submit work that you believe to be
  correct, and you will get significantly more partial credit if you
  clearly identify the gap(s) in your solution. It is good practice to
  start any long solution with an informal (but accurate) summary that
  describes the main idea. You may opt for the ``I take 15\%'' option.

\item You need to submit a PDF file before the deadline. Either a
  clear scan of you handwriting or a typeset document is accepted. You
  will get $\texbp$ bonus points for typing in LaTeX (Download and use
  the accompany TeX file).

\item You may collaborate with others on this problem set. However,
  you must \textbf{{write up your own solutions}} and \textbf{{list
      your collaborators and any external sources}} for each
  problem. Be ready to explain your solutions orally to a course staff
  if asked.

\item For problems that require you to provide an algorithm, you must
  give a precise description of the algorithm, together with a proof
  of correctness and an analysis of its running time. You may use
  algorithms from class as subroutines. You may also use any facts
  that we proved in class or from the book.

\item \textbf{If you describe a Greedy algorithm, you will get no
    credit without a formal proof of correctness, even if your
    algorithm is correct.}

\end{itemize}
\paragraph{Exercises. Do not turn in.}
\begin{questions}
\question Let $f$ and $f'$ be feasible $(s,t)$-flows in a flow network
  $G$, such that $v(f') > v(f)$. Prove that there is a feasible
  $(s,t)$-flow with value $v(f') - v(f)$ in the residual network
  $G_f$.

\question Let $u\to v$ be an arbitrary edge in an arbitrary flow
  network $G$. Prove that if there is a minimum $(s,t)$-cut $(S,T)$
  such that $u\in S$ and $v\in T$, then there is \emph{no} minimum cut
  $(S',T')$ such that $u\in T'$ and $v\in S'$.

  \question Describe a linear program for the bipartite maximum
  matching problem. The input is a bipartite graph
  $G = (U \cup V; E)$, where $E \subseteq U \times V$; the output is
  the largest matching in $G$. Your linear program should have one
  variable for each edge.

\end{questions}
\newpage 
\paragraph{Problems to turn in.}

\begin{questions}
  \question[10] (Stabbing points) Let $X$ be a set of $n$ intervals on
  the real line. We say that a set $P$ of points \emph{stabs} $X$ if
  every interval in $X$ contains at least one point in $P$. Describe
  an efficient algorithm to compute the smallest set of points that
  stabs $X$. Assume that your input consists of two arrays
  $L[1,\ldots,n]$ and $R[1,\ldots,n]$, representing the left and right
  endpoints of the intervals in $X$.

  \newpage

  \question[10] (Maximum spanning tree) Describe an algorithm to
  compute the \emph{maximum}-weight spanning tree of a given
  edge-weighted graph.

  \newpage

  \question[10] Suppose we are maintaining a data structure under a
  series of $n$ operations. Let $f(i)$ denote the actual running time
  of the $i$th operation. Suppose that $f(i)$ is the largest integer
  $k$ such that $2^k$ divides $i$. Determine the resulting amortized
  cost of a single operation.
  
\newpage 
\question[15] (Vertex cover) A \emph{vertex cover} of an undirected
  graph $G = (V, E)$ is a subset of the vertices which touches every
  edge—that is, a subset $S \subseteq V$ such that for each edge
  ${u,v} \in E$, one or both of $u,v$ are in $S$.  Describe and
  analyze an algorithm, as efficient as you can, to find a minimum
  vertex cover in a bipartite graph.

  \newpage

  \question (Updating max flow) You are given a flow network
  $G = (V,E)$ with source $s$ and sink $t$, and integer capacities.
  \begin{parts}
    \part[10] Suppose that you are given a max flow in $G$. Now we
    increase the capacity of a single edge $(u,v)\in E$ by 1. Give an
    $O(m+n)$-time algorithm to update the max flow.
    \part[10] Now suppose all edges have unit capacity and you are
    given a parameter $k$. The goal is to delete $k$ edges so as to
    reduce the maximum $s-t$ flow in $G$ as much as possible. In other
    words, you should find a set of edges $F\subseteq E$ so that
    $|F|=k$ and the maximum $s-t$ flow in $G'=(V,E-F)$ is as small as
    possible subject to this. Describe a polynomial-time algorithm to
    solve this problem.
  \end{parts}
\end{questions}
\end{document}


